### 2016-05-10 capability-machine.tex
Added a standard update of pc to the semantics
Updated ticket machine example to use restrict instruction with correct number of args.
Added the remaining cases missing in the V-relation
Wrote an initial attempt for a lemma about the ticket dispenser.

\newcommand{\stdcap}[1][\perm]{\left(#1,\start,\addrend,\offset \right)}

@@

%Functions
\newcommand{\plainfun}[1]{\mathit{#1}}
\newcommand{\decode}{\plainfun{decode}}
\newcommand{\encode}{\plainfun{encode}}
\newcommand{\encodePerm}{\plainfun{encodePerm}}
\newcommand{\updatePcPerm}[1]{\plainfun{updatePcPerm}(#1)}
\newcommand{\executeAllowed}[1]{\plainfun{executeAllowed}(#1)}
\newcommand{\nonZero}[1]{\plainfun{nonZero}(#1)}
\newcommand{\readAllowed}[1]{\plainfun{readAllowed}(#1)}
\newcommand{\writeAllowed}[1]{\plainfun{writeAllowed}(#1)}
\newcommand{\withinBounds}[1]{\plainfun{withinBounds}(#1)}

@@

@@

                                               \updateReg{\pcreg}{\memreg(\pcreg) + 1} & \text{if not $\nonZero{\memreg(\rv)}$}\\

@@

                                               \updateReg{r_1}{\var{w}} &

@@

                                               \updateHeap{\addr}{\var{w}} &

@@

                                                 \updateReg{r_1}{\memreg(\rv)} & \text{if $r_1 \neq \pcreg$} \\ % For now we do this, so you need to use a jump to move stuff into pc.

@@

                                               \updateReg{r_1}{\var{c}} &

@@

                                                            \updateReg{r_1}{n_1+n_2} &

@@

                                                              \qquad\text{$n_i = \rv_i$ or $n_i = \memreg{\rv_i}$}

@@

\section{Examples}
\label{sec:examples}
\subsection{Ticket Dispenser}
\newcommand{\size}{\var{size}}
\label{sec:tick-disp}
\[
  \begin{array}{r l}
    \init: & \move{\refreg{r_1}}{\size} \\
           & \move{\refreg{r_0}}{\refreg{\pcreg}} \\
           & \lea{\refreg{r_0}}{3} \\
           & \jmp{\malloc} \\
           & \store{\refheap{r_1}}{(\encode(i_1))} \\
           & \lea{\refreg{r_1}}{1} \\
           & \store{\refheap{r_1}}{(\encode(i_2))} \\
           & \lea{\refreg{r_1}}{1} \\
           & \vdots \\
           & \store{\refheap{r_1}}{(\encode(i_6))} \\
           & \lea{\refreg{r_1}}{1} \\
           & \move{\refreg{r_0}}{\refreg{r_1}} \\
           & \lea{\refreg{r_0}}{1} \\
           & \store{\refheap{r_1}}{\refreg{r_0}} \\
           & \lea{\refreg{r_1}}{1} \\
           & \store{\refheap{r_1}}{0} \\
           & \lea{\refreg{r_1}}{-7} \\
           & \restrict{\refreg{r_1}}{(\encodePerm{(\entry)})} \\
           & \move{\refreg{r_0}}{\refreg{\pcreg}} \\
           & \lea{\refreg{r_0}}{4} \\
           & \restrict{\refreg{r_0}}{(\encodePerm{(\entry)})} \\
           & \jmp{\codelabel{adv}}
  \end{array}
\]
Here $\size$ is 9. $i_1 \dots i_7$ refers to the instructions in the following program:
\[
  \begin{array}{r l}
    i_1 :& \move{\refreg{r_2}}{\refreg{\pcreg}} \\
    i_2 :& \load{\refreg{r_2}}{\size - 2} \\
    i_3 :& \load{\refreg{r_1}}{\refheap{r_2}}\\
    i_4 :& \plus{\refreg{r_3}}{\refreg{r_1}}{2} \\
    i_5 :& \store{\refheap{r_2}}{\refreg{r_3}} \\
    i_6 :& \move{\refreg{r_2}}{0} \\
    i_7 :& \jmp{\refreg{r_0}}
  \end{array}
\]
The heap layout for the above program can be seen in Figure~\ref{tab:tick-disp-heap}.
\begin{figure}[h]
  \centering
  \begin{tabular}{ | c | l | }
    \hline
    Addr. & Content \\ \hline
    0   &  $i_1$ \\ \hline
    1   &  $i_2$ \\ \hline
    2   &  $i_3$ \\ \hline
    3   &  $i_4$ \\ \hline
    4   &  $i_5$ \\ \hline
    5   &  $i_6$ \\ \hline
    5   &  $i_7$ \\ \hline
    7   & Cap. for addr. 8 \\ \hline
    8   & The counter \\ \hline  
  \end{tabular}
  \caption{The heap layout for the ticket dispenser.}
  \label{tab:tick-disp-heap}
\end{figure}

@@

\[
\stdvr \; : \;  \Worlds \monfun \UPred{\Words}
\]
\begin{align*}
  \stdvr(W) \defeq & \{ \npair{i} \mid i \in \ints \} 
\union \\
                   & \{ \npair{\stdcap[\noperm] } \} 
\union \\
                   & \{ \npair{\stdcap[\readonly]} \mid \\
                   & \quad \exists r \in \RegionName \ldotp \\
                   & \qquad  \forall \addr \in \{\start, \dots, \addrend \} \ldotp \\
                   & \qquad \quad  \forall \npair{\heapseg} \in W(r)(W)\ldotp \\
                   & \qquad \qquad   \addr \in \dom(\heapseg) \land \\
                   & \qquad \qquad   \npair[n-1]{\heapseg(\addr)} \in \stdvr(W)\} 
\union \\
                   & \{ \npair{\readwrite} \mid \\
                   & \quad \exists r \in \RegionName \ldotp \\
                   & \qquad  \forall \addr \in \{\start, \dots, \addrend \} \ldotp \\
                   & \qquad \quad  \forall \npair{\heapseg} \in W(r)(W)\ldotp \\
                   & \qquad \qquad   \addr \in \dom(\heapseg) \land \\
                   & \qquad \qquad   \npair[n-1]{\heapseg(\addr)} \in \stdvr(W) \land \\
                   & \qquad \qquad   \forall n' < n, W' \future W, v \ldotp \\
                   & \qquad \qquad \quad \heapSat{n'}{W'} \land \\
                   & \qquad \qquad \quad \npair[n']{v} \in \stdvr(W') \\
                   & \qquad \qquad \qquad  \Rightarrow \exists W'' \future W' \ldotp \heapSat[\memheap\update{a}{v}]{n''}{W''} \}
\union \\
                   & \{ \npair{\stdcap[\exec] = c} \mid \\
                   & \quad (\exists r \in \RegionName \ldotp \\
                   & \qquad  \forall \addr \in \{\start, \dots, \addrend \} \ldotp \\
                   & \qquad \quad  \forall \npair{\heapseg} \in W(r)(W)\ldotp \\
                   & \qquad \qquad   \addr \in \dom(\heapseg) \land \\
                   & \qquad \qquad   \npair[n-1]{\heapseg(\addr)} \in \stdvr(W)) \\
                   & \quad \land \\
                   & \quad \forall n' < n \ldotp \forall W' \future W \ldotp \\
                   & \qquad \forall \npair[n']{v_1,\dots,v_k} \in \stdvr(W') \ldotp \\
                   & \qquad \quad \npair[n']{[\pcreg \mapsto c, r_1 \mapsto v_1, \dots , r_k \mapsto v_k]} \in \stder(W') \}
\}
\end{align*}

>>>>

\newcommand{\stdcap}[1][\perm]{\left(#1,\start,\addrend,\addr \right)}

@@

%Functions
\newcommand{\plainfun}[1]{\mathit{#1}}
\newcommand{\decode}{\plainfun{decode}}
\newcommand{\encode}{\plainfun{encode}}
\newcommand{\encodePerm}{\plainfun{encodePerm}}
\newcommand{\updatePcPerm}[1]{\plainfun{updatePcPerm}(#1)}
\newcommand{\executeAllowed}[1]{\plainfun{executeAllowed}(#1)}
\newcommand{\nonZero}[1]{\plainfun{nonZero}(#1)}
\newcommand{\readAllowed}[1]{\plainfun{readAllowed}(#1)}
\newcommand{\writeAllowed}[1]{\plainfun{writeAllowed}(#1)}
\newcommand{\withinBounds}[1]{\plainfun{withinBounds}(#1)}
\newcommand{\stdUpdatePc}[1]{\plainfun{updatePc(#1)}}

@@

  \stdUpdatePc{\Phi} &=
                         \updateReg{\pcreg}{\var{newPc}} & 
                             \quad\text{and $\var{newPc} = (\perm,\start,\addrend,\addr + 1)$}\\

@@

                                               \stdUpdatePc{\Phi} & \text{if not $\nonZero{\memreg(\rv)}$}\\

@@

                                               \stdUpdatePc{\updateReg{r_1}{\var{w}}} &

@@

                                               \stdUpdatePc{\updateHeap{\addr}{\var{w}}} &

@@

                                                 \stdUpdatePc{\updateReg{r_1}{\memreg(\rv)}} & \text{if $r_1 \neq \pcreg$} \\ % For now we do this, so you need to use a jump to move stuff into pc.

@@

                                               \stdUpdatePc{\updateReg{r_1}{\var{c}}} &

@@

                                                            \stdUpdatePc{\updateReg{r_1}{n_1+n_2}} &

@@

                                                              \qquad\text{$n_i = \rv_i$ or $n_i = \memreg{\rv_i}$}\\

@@

\section{Examples}
\label{sec:examples}
\subsection{Ticket Dispenser}
\newcommand{\size}{\var{size}}
Assume the instructions of some adversary resides in the memory starting at the memory location marked with $\codelabel{adv}$.
\label{sec:tick-disp}
\[
  \begin{array}{r l}
    \init: & \move{\refreg{r_1}}{\size} \\
           & \move{\refreg{r_0}}{\refreg{\pcreg}} \\
           & \lea{\refreg{r_0}}{3} \\
           & \jmp{\malloc} \\
           & \store{\refheap{r_1}}{(\encode(i_1))} \\
           & \lea{\refreg{r_1}}{1} \\
           & \store{\refheap{r_1}}{(\encode(i_2))} \\
           & \lea{\refreg{r_1}}{1} \\
           & \vdots \\
           & \store{\refheap{r_1}}{(\encode(i_6))} \\
           & \lea{\refreg{r_1}}{1} \\
           & \move{\refreg{r_0}}{\refreg{r_1}} \\
           & \lea{\refreg{r_0}}{1} \\
           & \store{\refheap{r_1}}{\refreg{r_0}} \\
           & \lea{\refreg{r_1}}{1} \\
           & \store{\refheap{r_1}}{0} \\
           & \lea{\refreg{r_1}}{-7} \\
           & \restrict{\refreg{r_1}}{\refreg{r_1}}{(\encodePerm{(\entry)})} \\
           & \move{\refreg{r_0}}{\refreg{\pcreg}} \\
           & \lea{\refreg{r_0}}{4} \\
           & \restrict{\refreg{r_0}}{\refreg{r_0}}{(\encodePerm{(\entry)})} \\
           & \jmp{\codelabel{adv}}
  \end{array}
\]
Here $\size$ is 9. $i_1 \dots i_7$ refers to the instructions in the following program:
\[
  \begin{array}{r l}
    i_1 :& \move{\refreg{r_2}}{\refreg{\pcreg}} \\
    i_2 :& \load{\refreg{r_2}}{\size - 2} \\
    i_3 :& \load{\refreg{r_1}}{\refheap{r_2}}\\
    i_4 :& \plus{\refreg{r_3}}{\refreg{r_1}}{2} \\
    i_5 :& \store{\refheap{r_2}}{\refreg{r_3}} \\
    i_6 :& \move{\refreg{r_2}}{0} \\
    i_7 :& \jmp{\refreg{r_0}}
  \end{array}
\]
The heap layout for the above program can be seen in Figure~\ref{tab:tick-disp-heap}.
\begin{figure}[h]
  \centering
  \begin{tabular}{ | c | l | }
    \hline
    Addr. & Content \\ \hline
    0   &  $i_1$ \\ \hline
    1   &  $i_2$ \\ \hline
    2   &  $i_3$ \\ \hline
    3   &  $i_4$ \\ \hline
    4   &  $i_5$ \\ \hline
    5   &  $i_6$ \\ \hline
    5   &  $i_7$ \\ \hline
    7   & Cap. for addr. 8 \\ \hline
    8   & The counter \\ \hline  
  \end{tabular}
  \caption{The heap layout for the ticket dispenser.}
  \label{tab:tick-disp-heap}
\end{figure}
In the JS paper, the ticket dispenser program dereferences the counter and returns the value. It does not seem like a similar thing would work here as the adversary can do even more. In particular, the adversary can execute a $\halt$ instruction that will cause the machine to halt (successfully). The adversary could also choose never to return to the return address, which we set up for him.
\begin{lemma}
 Given $\Phi$, if $\memheap$ contains the ticket dispenser program starting at the address labelled $\init$, $\memreg(\pcreg) = (\exec,\init,\init+26,\init)$, and $\Phi \rightarrow^* (\halted,\Phi')$, then if $i$ was the address returned by the call to $\malloc$, then $\refheap(i+8)$ is even and positive.
\end{lemma}
The magic number 26 is the size of the ticet dispenser program (minus one), so the $\Phi$ is assumed to have an execute permission for the area of the heap where the program resides.

@@

\[
\stdvr \; : \;  \Worlds \monfun \UPred{\Words}
\]
  \stdvr(W) \defeq & \{ \npair{i} \mid i \in \ints \} 
\union \\
                   & \{ \npair{\stdcap[\noperm] } \} 
\union \\
                   & \{ \npair{\stdcap[\readonly]} \mid \\
                   & \quad \exists r \in \RegionName \ldotp \\
                   & \qquad  \forall \addr \in \{\start, \dots, \addrend \} \ldotp \\
                   & \qquad \quad  \forall \npair{\heapseg} \in W(r)(W)\ldotp \\
                   & \qquad \qquad   \addr \in \dom(\heapseg) \land \\
                   & \qquad \qquad   \npair[n-1]{\heapseg(\addr)} \in \stdvr(W)\} 
\union \\
                   & \{ \npair{\readwrite} \mid \\
                   & \quad \exists r \in \RegionName \ldotp \\
                   & \qquad  \forall \addr \in \{\start, \dots, \addrend \} \ldotp \\
                   & \qquad \quad  \forall \npair{\heapseg} \in W(r)(W)\ldotp \\
                   & \qquad \qquad   \addr \in \dom(\heapseg) \land \\
                   & \qquad \qquad   \npair[n-1]{\heapseg(\addr)} \in \stdvr(W) \land \\
                   & \qquad \qquad   \forall n' < n, W' \future W, v \ldotp \\
                   & \qquad \qquad \quad \heapSat{n'}{W'} \land \\
                   & \qquad \qquad \quad \npair[n']{v} \in \stdvr(W') \\
                   & \qquad \qquad \qquad  \Rightarrow \exists W'' \future W' \ldotp \heapSat[\memheap\update{a}{v}]{n''}{W''} \}
\union \\ & \dots\text{Continues on next page.}
                   & \{ \npair{\stdcap[\exec] = c} \mid \\
                   & \quad (\exists r \in \RegionName \ldotp \\
                   & \qquad  \forall \addr \in \{\start, \dots, \addrend \} \ldotp \\
                   & \qquad \quad  \forall \npair{\heapseg} \in W(r)(W)\ldotp \\
                   & \qquad \qquad   \addr \in \dom(\heapseg) \land \\
                   & \qquad \qquad   \npair[n-1]{\heapseg(\addr)} \in \stdvr(W)) \\
                   & \quad \land \\
                   & \quad \forall n' < n \ldotp \forall W' \future W \ldotp \\
                   & \qquad \forall \npair[n']{v_1,\dots,v_k} \in \stdvr(W') \ldotp \\
                   & \qquad \quad \npair[n']{[\pcreg \mapsto c, r_1 \mapsto v_1, \dots , r_k \mapsto v_k]} \in \stder(W') \}
\union \\
% For the entry case a capability is acceptable, if we can change it to an execute permission and put it as the pc and we otherwise have valid words in the registers, then the register file should be valid.
                   & \{ \npair{\stdcap[\entry]} \mid \\
                   & \quad \forall n' < n \ldotp \forall W' \future W \ldotp \\
                   & \qquad \forall \npair[n']{v_1,\dots,v_k} \in \stdvr(W') \ldotp \\
                   & \qquad \quad \npair[n']{[\pcreg \mapsto (\stdcap[\exec]), r_1 \mapsto v_1, \dots , r_k \mapsto v_k]} \in \stder(W') \}
\union \\
                   & \{ \npair{\stdcap[\exec] = c} \mid \\
                   & \quad \exists r \in \RegionName \ldotp \\
                   & \qquad  \forall \addr \in \{\start, \dots, \addrend \} \ldotp \\
                   & \qquad \quad  \forall \npair{\heapseg} \in W(r)(W)\ldotp \\
                   & \qquad \qquad   \addr \in \dom(\heapseg) \land \\
                   & \qquad \qquad   \npair[n-1]{\heapseg(\addr)} \in \stdvr(W) \land \\
                   & \qquad \qquad   \forall n' < n, W' \future W, v \ldotp \\
                   & \qquad \qquad \quad \heapSat{n'}{W'} \land \\
                   & \qquad \qquad \quad \npair[n']{v} \in \stdvr(W') \\
                   & \qquad \qquad \qquad  \Rightarrow \exists W'' \future W' \ldotp \heapSat[\memheap\update{a}{v}]{n''}{W''} \}
                   & \quad \land \\
                   & \quad \forall n' < n \ldotp \forall W' \future W \ldotp \\
                   & \qquad \forall \npair[n']{v_1,\dots,v_k} \in \stdvr(W') \ldotp \\
                   & \qquad \quad \npair[n']{[\pcreg \mapsto c, r_1 \mapsto v_1, \dots , r_k \mapsto v_k]} \in \stder(W') \}
\}
